Tropical mathematics and the Lambda-calculus I

### 2023-08-02 CSL24/section4.tex
Commit

As seen in Section \ref{section2}, tropical polynomials are piecewise-linear functions, hence concave and Lipschitz-continuous. Moreover, tps in finitely many variables are locally equivalent to tropical polynomials (except at some singular points), and are thus also concave and locally Lipschitz-continuous.
In this section we show that much of these properties extend also to tps with infinitely many variables, as those arising from the tropical relational model. %establish some topological and metric properties of tps, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
The literature on tropical power series is often recent (e.g.~\cite{Porzio2021}), and several results we prove in this section are, to our knowledge, new.
%
%Remark that the map $(\_)^!$ is a surjective homomorphism (w.r.t.\ coKleisli composition and identity and function composition and identiy) on its image. {\color{red}In Ehrhard-Danos fanno vedere (Lemma 14) che \`e anche iniettivo. Non riesco a capire se anche a noi lo \`e, ma non vedo perch\'e non dovrebbe.}
%Remark that, since $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}\in\HOM{\LREL_!}{\mathlarger{\&}_{i=1}^n \model{A_i}}{\model B}\simeq\Lawv^{\prod_{i=1}^n ! \model {A_i}\times \model B}$, the tps $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}^!:\Lawv^{\mathlarger{+}_{i=1}^n \model{A_i}}\to\Lawv^{\model B} \simeq \prod_{i=1}^n\Lawv^{\model{A_i}}\to\Lawv^{\model B}$ has shape $t^!(x^1,\dots,x^n)_b=\inf_{\mu_i\in ! \model{A_i}}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Same for $\STDLC$.
%
%\begin{remark}
%The interpretations of $\BSTLC$-terms are matrices $t\in\HOM{\LREL}{\bigotimes_{i=1}^n !_{n_i}X_i}{Y}=\Lawv^{\prod_{i=1}^n !_{n_i}X_i\times Y}$.
%In such situation, we define $t^!:\Lawv^{\mathlarger{+}_{i=1}^n X_{i}}\to\Lawv^Y \simeq \prod_{i=1}^n\Lawv^{X_{i}}\to\Lawv^Y$ as $t^!(x^1,\dots,x^n)_b:=\inf_{\mu_i\in !_{n_i} X_i}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Clearly $t^!={\widetilde t}^!$, where $\widetilde t\in\Lawv^{\prod_{i=1}^n ! {X_i}\times Y}$ is the matrix $\widetilde t_{(\mu_1,\dots,\mu_n),b}:=t_{(\mu_1,\dots,\mu_n),b}$ if $\mu_i\in !_{n_i} X_i$ for all $i$, and $:=\infty$ otherwise, which has the same support as $t$.
%So, such $t^!$'s are special cases of tps and thus we can \emph{Wlog} only consider tps's.
%\end{remark}

@@

%
%%In Example~\ref{ex:famous_ex}, 
%Again by looking at Fig~\ref{fig:plot1} it appears that, \emph{far from $0$}, $\varphi$ behaves like some polynomials $\varphi_{n}$.
%In particular, %for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. More, precisely, 
%$\varphi$ coincides on $[\epsilon,\infty]$ with $\varphi_{n}$,
%for $\epsilon \geq 2^{-(n+1)}$ (the smallest tropical root of $\varphi_{n}$).
%However, at
%%
%%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
% $x=0$ we have that $\varphi(x=0)=\inf_{n\in\N} 2^{-n}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
%Also, while the derivative of $f$ is bounded on all $(0,\infty)$, for $x\to 0^+$ it tends to $\infty$.
%This phenomenon is reminiscent of \cite[Example 7]{Ehrhard2005},
%%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019
%which actually motivated our first investigations.
%In fact, this behaviour is shared by all tps with \emph{finitely} many variables, as shown by the following result (we identify $!\set{1,\dots,k}\simeq \N^k$, so the matrix of a tps $f$ with finitely many variables $x=x_1,\dots,x_k$ (and one output variable) is given as a $\matr f:\N^k\to\Lawv$, $f$ having shape $f(x)=\inf_{n\in \N^k}\set{nx+\matr f(n)}$, with $nx$ the scalar product).
%
%\begin{theorem}\label{theorem:fepsilon}
% Let $k\in\N$ and $f:\Lawv^k\to\Lawv$ a tps with matrix $\matr f:\N^k\to\Lawv$.
% For all $0<\epsilon<\infty$, there is a \emph{finite} $\C F_\epsilon \subseteq \N^k$ such that 
%% 
%% \begin{enumerate}
%%  \item If $\C F_\epsilon=\emptyset$ then $f=\infty$ on all $\Lawv^k$;
%%  \item If $f(x_0)=\infty$ for some $x_0\in[0,\infty)^k$, then $\C F_\epsilon=\emptyset$;
%  %\item 
%$f$ coincides on all $[\epsilon,\infty]^k$ with the tropical \emph{polynomial} $P_\epsilon(x):=\min_{n\in \C F_\epsilon}\set{nx+\matr f(n)}$.
%% \end{enumerate}
%\end{theorem}
%\begin{proof}[Proof sketch]
%Let $\C F_\epsilon$ be the set of $n\in\N^{k}$ such that 
%$\widehat f(n)<\infty$ and $\widehat f(m)> \widehat f(n)+\epsilon$ holds for all $m\prec n$, where $\preceq$ is the pointwise order on $\N^k$.
%The core of the proof is showing that this set is indeed finite and enough for computing $f$.
%\end{proof}
%
In the remainder of this section, by a tropical polynomial or power series we mean one with possibly infintely many variables. Moreover, we consider the sets $\Lawv^X$ endowed with the usual sup-norm $\norm{x}_\infty:=\sup_{a\in X} x_a$.

@@

\begin{proposition}\label{prop:nondecr+conc}
 All tps 
 %$f:\Lawv^X\to\Lawv^Y$
are non-decreasing and concave, w.r.t.\ the pointwise order on $\Lawv^{X}$.
\end{proposition}

@@

\begin{theorem}\label{thm:cont}
 All tps $f:\Lawv^X\to\Lawv$ are continuous on $(0,\infty)^X$, w.r.t.\ to the norm $\norm{\cdot}_\infty$.
\end{theorem}
\begin{proof}
%It follows after adapting \cite[Proposition 4.4]{Cobzas2017} in order to prove that if a real-valued function on a locally convex topological $\BB R$-vector space is, locally around $x$, concave and bounded by a finite constant, then it is continuous at $x$.
%Now, our $f$ is concave, so 
By \autoref{prop:nondecr+conc}, $-f$ is convex.
Since $f\geq 0$ on all $\Lawv^X$, we have e.g.\ $-f\leq 1$ on $\Lawv^X$.
%Remark that we cannot immediately apply \cite[Proposition 4.4.(3)]{Cobzas2017}, since $\Lawv^X$ is not even a $\BB R$-vector space.
Now $(0,\infty)^X\subseteq \BB R^X$ is open and convex, %and {\color{red}$\BB R^X$ is a LCTVS}, 
so \cite[Proposition 4.4.(3)]{Cobzas2017} entails the continuity of $-f$ on $(0,\infty)^X$, hence that of $f$ on it.
\end{proof}

We conclude this subsection by noticing that, as a set, $\Lawv=[0,\infty]$, and  $[0,\infty]^X$ with the usual $+$ and $\cdot$ is a $\BB R_{\geq0}$-semimodule, let us call it $\overline{\BB R}_{\geq 0}^X$.
Together with the norm $\norm{\cdot}_\infty$, it can be showed to be a Scott-complete \emph{normed cone} (see~\cite{Selinger2004} or the appendix).
Its cone structure induces an order on it, called the \emph{cone order}:
$x\leq y$ iff $y=x+z$ for some (unique) $z\in\overline{\BB R}_{\geq 0}^X$.
Such order actually coincides with the pointwise order on $\overline{\BB R}_{\geq 0}$.
This makes it a Scott-continuous dcpo.
Suitable categories of cones have been recently investigated as models of probabilistic computation (\cite{Crubillie2018, EhrPagTas2018, Ehrhard2020}).
In analogy with \cite[Proposition 17]{DanEhrh2011}, we have:

@@

\begin{proposition}\label{prop:troplinear}
\begin{enumerate}
\item If a tps $f: \Lawv^{X}\to \Lawv^{Y}$ arises from a matrix $\matr f:X\times Y\to \Lawv$ (i.e.~it is tropical linear), then $f$ is non-expansive (i.e.\ $1$-Lipschitz).  
\item If  $f: \Lawv^{X}\to \Lawv^{Y}$ arises from a matrix $\matr f: \ !_{n}X\times Y\to \Lawv$, then $f$ is $n$-Lipschitz-continuous.
\end{enumerate}
\end{proposition}
\begin{proof}[Proof sketch]
1). Using the fact that $f( x)_{b}= \inf_{a\in X}\{\matr f_{a,b}+ x_{a}\}$,
the problem reduces to checking that $|(\matr f_{a,b}- x_{a})- (\matr f_{a,b}- y_{a})| = |x_{a}- y_{a}|\leq \|  x- y\|_{\infty}$ (see the Appendix).

@@

Observe that, for any tropical polynomial $\varphi:\Lawv^{X}\to \Lawv^{Y}$, the associated matrix has shape $!_{\mathrm{deg}(\varphi)}(X)\times Y\to \Lawv$ (as a monomial $\mu_ix+c_{i}$ yields a matrix entry on $!_{\#\mu_i}X\times Y$). Hence, using Proposition \ref{prop:troplinear} 2., we have:
\begin{corollary}\label{prop:polylip}
For any tropical polynomial $\varphi:\Lawv^{X}\to\Lawv$, $\varphi$ is $\mathrm{deg}(\varphi)$-Lipschitz continuous.
\end{corollary}

@@

\begin{theorem}\label{thmTLSlocLip}
%Let $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $d$ any metric.
%If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
%Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
%Therefore, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, %and all tps are non-negative, 
%we immediately have that 
All tps $f:\Lawv^X\to\Lawv$ are locally Lipschitz on $(0,\infty)^X$.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} f$.
\end{theorem}
\begin{proof}
By observing that $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, we apply the result that for all $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$ concave and locally bounded, with $V$ open and convex and $d$ any metric, $f$ is locally Lipschitz, with the stated Lipschitz constant on $\overline{B_{\delta}(x)}$. %
% Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
\end{proof}
%\begin{proof}[Proof]
% Call $\overline{B_{\delta}(x)}:=B_1$, $\overline{B_{3\delta}(x)}:=B_3$.
% It suffices to show that for all $x\in V$, there is $\delta>0$ s.t.\ $B_3\subseteq \mathrm{interior}(V)$, $K:=\max_{B_3}  \absv f$ exists and $f_{\big|_{B_1}}$ is $(\frac{1}{\delta}\max_{B_3} \absv f)$-Lipschitz.
%A $\delta$ satisfying the first two conditions exists since $V$ is open and bcause $f$ is locally bounded and $B_3$ is compact.
%Now fix $y,z\in B_1$.
%Call $r:=\frac{d(y,z)}{2\delta}\in[0,1]$.
%We have to show that $\absv{f(y)-f(z)}\leq \frac{K}{\delta}d(y,z)=2Kr$.
%Wlog $y\neq z$.
%So $r\neq 0$ and we can consider $u:=\frac{1+r}{r}z-\frac{1}{r}y$, $v:=\frac{1}{r}y-\frac{r-1}{r}z$.
%We have $u,v\in \overline{B_{2\delta}(z)}=:B_2$ (geometrically, those are the intersections between $B_2$ and the segment between $y$ and $z$).
%Now we have the convex combinations $z=\frac{1}{1+r}y+\frac{r}{1+r}u$ and $y=(1-r)z+rv$, so the concavity of $f$ entails on one hand:
%$f(z)\geq \frac{1}{1+r}f(y)+\frac{r}{1+r}f(z)\geq \frac{f(y)}{1+r} - \frac{rK}{1+r}$, i.e.\ $f(y)-f(z)\leq r(K+f(z))\leq 2rK$, and on the other hand:
%$f(y)\geq (1-r)f(z)+rf(v)\geq f(z)-r(f(z)+K)$, i.e.\ $f(z)-f(y)\leq r(f(z)+K)\leq 2rK$.
%The last inequalities follow because $B_2\subseteq B_3$.
%Putting those together, we have the thesis.
%\end{proof}

>>>>

As seen in Section \ref{section2}, tropical polynomials are piecewise-linear functions, hence concave and Lipschitz-continuous. Moreover, tps in finitely many variables are locally equivalent to tropical polynomials (except at some singular points), and are thus also concave and locally Lipschitz-continuous.
In this section we show that much of these properties extend also to tps with infinitely many variables, as those arising from the tropical relational model, %establish some topological and metric properties of tps, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
and by a tropical polynomial or power series we mean one with possibly infintely many variables.
The literature on tropical power series is often recent (e.g.~\cite{Porzio2021}), and several results we prove in this section are, to our knowledge, new.
%
%Remark that the map $(\_)^!$ is a surjective homomorphism (w.r.t.\ coKleisli composition and identity and function composition and identiy) on its image. {\color{red}In Ehrhard-Danos fanno vedere (Lemma 14) che \`e anche iniettivo. Non riesco a capire se anche a noi lo \`e, ma non vedo perch\'e non dovrebbe.}
%Remark that, since $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}\in\HOM{\LREL_!}{\mathlarger{\&}_{i=1}^n \model{A_i}}{\model B}\simeq\Lawv^{\prod_{i=1}^n ! \model {A_i}\times \model B}$, the tps $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}^!:\Lawv^{\mathlarger{+}_{i=1}^n \model{A_i}}\to\Lawv^{\model B} \simeq \prod_{i=1}^n\Lawv^{\model{A_i}}\to\Lawv^{\model B}$ has shape $t^!(x^1,\dots,x^n)_b=\inf_{\mu_i\in ! \model{A_i}}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Same for $\STDLC$.
%
%\begin{remark}
%The interpretations of $\BSTLC$-terms are matrices $t\in\HOM{\LREL}{\bigotimes_{i=1}^n !_{n_i}X_i}{Y}=\Lawv^{\prod_{i=1}^n !_{n_i}X_i\times Y}$.
%In such situation, we define $t^!:\Lawv^{\mathlarger{+}_{i=1}^n X_{i}}\to\Lawv^Y \simeq \prod_{i=1}^n\Lawv^{X_{i}}\to\Lawv^Y$ as $t^!(x^1,\dots,x^n)_b:=\inf_{\mu_i\in !_{n_i} X_i}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
%Clearly $t^!={\widetilde t}^!$, where $\widetilde t\in\Lawv^{\prod_{i=1}^n ! {X_i}\times Y}$ is the matrix $\widetilde t_{(\mu_1,\dots,\mu_n),b}:=t_{(\mu_1,\dots,\mu_n),b}$ if $\mu_i\in !_{n_i} X_i$ for all $i$, and $:=\infty$ otherwise, which has the same support as $t$.
%So, such $t^!$'s are special cases of tps and thus we can \emph{Wlog} only consider tps's.
%\end{remark}

@@

%
%%In Example~\ref{ex:famous_ex}, 
%Again by looking at Fig~\ref{fig:plot1} it appears that, \emph{far from $0$}, $\varphi$ behaves like some polynomials $\varphi_{n}$.
%In particular, %for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. More, precisely, 
%$\varphi$ coincides on $[\epsilon,\infty]$ with $\varphi_{n}$,
%for $\epsilon \geq 2^{-(n+1)}$ (the smallest tropical root of $\varphi_{n}$).
%However, at
%%
%%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
% $x=0$ we have that $\varphi(x=0)=\inf_{n\in\N} 2^{-n}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
%Also, while the derivative of $f$ is bounded on all $(0,\infty)$, for $x\to 0^+$ it tends to $\infty$.
%This phenomenon is reminiscent of \cite[Example 7]{Ehrhard2005},
%%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019
%which actually motivated our first investigations.
%In fact, this behaviour is shared by all tps with \emph{finitely} many variables, as shown by the following result (we identify $!\set{1,\dots,k}\simeq \N^k$, so the matrix of a tps $f$ with finitely many variables $x=x_1,\dots,x_k$ (and one output variable) is given as a $\matr f:\N^k\to\Lawv$, $f$ having shape $f(x)=\inf_{n\in \N^k}\set{nx+\matr f(n)}$, with $nx$ the scalar product).
%
%\begin{theorem}\label{theorem:fepsilon}
% Let $k\in\N$ and $f:\Lawv^k\to\Lawv$ a tps with matrix $\matr f:\N^k\to\Lawv$.
% For all $0<\epsilon<\infty$, there is a \emph{finite} $\C F_\epsilon \subseteq \N^k$ such that 
%% 
%% \begin{enumerate}
%%  \item If $\C F_\epsilon=\emptyset$ then $f=\infty$ on all $\Lawv^k$;
%%  \item If $f(x_0)=\infty$ for some $x_0\in[0,\infty)^k$, then $\C F_\epsilon=\emptyset$;
%  %\item 
%$f$ coincides on all $[\epsilon,\infty]^k$ with the tropical \emph{polynomial} $P_\epsilon(x):=\min_{n\in \C F_\epsilon}\set{nx+\matr f(n)}$.
%% \end{enumerate}
%\end{theorem}
%\begin{proof}[Proof sketch]
%Let $\C F_\epsilon$ be the set of $n\in\N^{k}$ such that 
%$\widehat f(n)<\infty$ and $\widehat f(m)> \widehat f(n)+\epsilon$ holds for all $m\prec n$, where $\preceq$ is the pointwise order on $\N^k$.
%The core of the proof is showing that this set is indeed finite and enough for computing $f$.
%\end{proof}
%
Notice that, as a set, $\Lawv^X=[0,\infty]^X$, and with the usual $+$ and $\cdot$ it is a $\BB R_{\geq0}$-semimodule, let us call it $\overline{\BB R}_{\geq 0}^X$.
Together with the usual sup-norm $\norm{x}_\infty:=\sup_{a\in X} x_a$, it can be showed to be a Scott-complete \emph{normed cone} (see~\cite{Selinger2004} or the appendix).
Suitable categories of cones have been recently investigated as models of probabilistic computation (\cite{Crubillie2018, EhrPagTas2018, Ehrhard2020}).
The cone structure of $\overline{\BB R}_{\geq 0}^X$ also induces a partial order on it, its \emph{cone partial-order}:
$x\leq y$ iff $y=x+z$ for some (unique) $z\in\overline{\BB R}_{\geq 0}^X$.
It actually coincides with the pointwise order on $\overline{\BB R}_{\geq 0}$ (and makes it a Scott-continuous dcpo).
In this section we consider tps w.r.t.\ this structure.
%Moreover, we consider the sets $\Lawv^X$ endowed with the usual sup-norm $\norm{x}_\infty:=\sup_{a\in X} x_a$.

@@

\begin{proposition}\label{prop:nondecr+conc}
 All tps 
 %$f:\Lawv^X\to\Lawv^Y$
are non-decreasing and concave, w.r.t.\ the pointwise order on $\overline{\BB R}_{\geq 0}^{X}$.
\end{proposition}

@@

\begin{theorem}\label{thm:cont}
 All tps $f:\overline{\BB R}_{\geq 0}^X\to\overline{\BB R}_{\geq 0}$ are continuous on $(0,\infty)^X$, w.r.t.\ to the norm $\norm{\cdot}_\infty$.
\end{theorem}
\begin{proof}
%It follows after adapting \cite[Proposition 4.4]{Cobzas2017} in order to prove that if a real-valued function on a locally convex topological $\BB R$-vector space is, locally around $x$, concave and bounded by a finite constant, then it is continuous at $x$.
%Now, our $f$ is concave, so 
By \autoref{prop:nondecr+conc}, $-f$ is convex.
Since $f\geq 0$ on all $\overline{\BB R}_{\geq 0}^X$, we have e.g.\ $-f\leq 1$ on $\overline{\BB R}_{\geq 0}^X$.
%Remark that we cannot immediately apply \cite[Proposition 4.4.(3)]{Cobzas2017}, since $\Lawv^X$ is not even a $\BB R$-vector space.
Now $(0,\infty)^X\subseteq \BB R^X$ is open and convex, %and {\color{red}$\BB R^X$ is a LCTVS}, 
so \cite[Proposition 4.4.(3)]{Cobzas2017} entails the continuity of $-f$ on $(0,\infty)^X$, hence that of $f$ on it.
\end{proof}

In analogy with \cite[Proposition 17]{DanEhrh2011}, we also have:

@@

\begin{proposition}\label{prop:troplinear}
\begin{enumerate}
\item If a tps $f: \overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$ arises from a matrix $\matr f:X\times Y\to \overline{\BB R}_{\geq 0}$ (i.e.~it is tropical linear), then $f$ is non-expansive (i.e.\ $1$-Lipschitz).  
\item If  $f: \overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$ arises from a matrix $\matr f: \ !_{n}X\times Y\to \overline{\BB R}_{\geq 0}$, then $f$ is $n$-Lipschitz-continuous.
\end{enumerate}
\end{proposition}
\begin{proof}[Proof sketch]
1). Using the fact that $f( x)_{b}= \inf_{a\in X}\{\matr f_{a,b}+ x_{a}\}$,
the problem reduces to checking that $|(\matr f_{a,b}- x_{a})- (\matr f_{a,b}- y_{a})| = |x_{a}- y_{a}|\leq \|  x- y\|_{\infty}$ (see the Appendix).

@@

Observe that, for any tropical polynomial $\varphi:\overline{\BB R}_{\geq 0}^{X}\to \overline{\BB R}_{\geq 0}^{Y}$, the associated matrix has shape $!_{\mathrm{deg}(\varphi)}(X)\times Y\to \overline{\BB R}_{\geq 0}$ (as a monomial $\mu_ix+c_{i}$ yields a matrix entry on $!_{\#\mu_i}X\times Y$). Hence, using Proposition \ref{prop:troplinear} 2., we have:
\begin{corollary}\label{prop:polylip}
Any tropical polynomial $\varphi:\overline{\BB R}_{\geq 0}^{X}\to\overline{\BB R}_{\geq 0}$ is $\mathrm{deg}(\varphi)$-Lipschitz continuous.
\end{corollary}

@@

\begin{theorem}\label{thmTLSlocLip}
%Let $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $d$ any metric.
%If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
%Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
%Therefore, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, %and all tps are non-negative, 
%we immediately have that 
All tps $f:\overline{\BB R}_{\geq 0}^X\to\overline{\BB R}_{\geq 0}$ are locally Lipschitz on $(0,\infty)^X$.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} f$.
\end{theorem}
\begin{proof}
By observing that $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, we apply the result that for all $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$ concave and locally bounded, with $V$ open and convex and $d$ any metric, $f$ is locally Lipschitz, with the stated Lipschitz constant on $\overline{B_{\delta}(x)}$. %
% Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
\end{proof}
%\begin{proof}[Proof]
% Call $\overline{B_{\delta}(x)}:=B_1$, $\overline{B_{3\delta}(x)}:=B_3$.
% It suffices to show that for all $x\in V$, there is $\delta>0$ s.t.\ $B_3\subseteq \mathrm{interior}(V)$, $K:=\max_{B_3}  \absv f$ exists and $f_{\big|_{B_1}}$ is $(\frac{1}{\delta}\max_{B_3} \absv f)$-Lipschitz.
%A $\delta$ satisfying the first two conditions exists since $V$ is open and bcause $f$ is locally bounded and $B_3$ is compact.
%Now fix $y,z\in B_1$.
%Call $r:=\frac{d(y,z)}{2\delta}\in[0,1]$.
%We have to show that $\absv{f(y)-f(z)}\leq \frac{K}{\delta}d(y,z)=2Kr$.
%Wlog $y\neq z$.
%So $r\neq 0$ and we can consider $u:=\frac{1+r}{r}z-\frac{1}{r}y$, $v:=\frac{1}{r}y-\frac{r-1}{r}z$.
%We have $u,v\in \overline{B_{2\delta}(z)}=:B_2$ (geometrically, those are the intersections between $B_2$ and the segment between $y$ and $z$).
%Now we have the convex combinations $z=\frac{1}{1+r}y+\frac{r}{1+r}u$ and $y=(1-r)z+rv$, so the concavity of $f$ entails on one hand:
%$f(z)\geq \frac{1}{1+r}f(y)+\frac{r}{1+r}f(z)\geq \frac{f(y)}{1+r} - \frac{rK}{1+r}$, i.e.\ $f(y)-f(z)\leq r(K+f(z))\leq 2rK$, and on the other hand:
%$f(y)\geq (1-r)f(z)+rf(v)\geq f(z)-r(f(z)+K)$, i.e.\ $f(z)-f(y)\leq r(f(z)+K)\leq 2rK$.
%The last inequalities follow because $B_2\subseteq B_3$.
%Putting those together, we have the thesis.
%\end{proof}
